The Effects of weather on maize yield: New evidence from Kenya

### 2019-01-15 writing/draft4/draft21.tex
writing, or more like editing

To verify our results, we adapted an alternative step-down model building
approach using the Satterthwaite’s method to determine the p-values of the individual
t-tests \cite{lmerTest}. We started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the first part of the procedure, insignificant random effects were removed one by one. In each iteration, the insignificant variable with the highest p-value was removed. This was repeated until only significant random effects remained. After this, we applied an analogous procedure to the fixed effects until we were left only with significant variables. As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}. This method led us to the same results as those obtained with our primary AIC-based method described above.

According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$). To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with an ARMA($p$,$q$) error structure. In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$. Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$). The value of the AIC criterion of the model with the ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.  The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $0.003$, hence the ARMA($1$,$1$) error structure turned out to be a better fit. The AIC criteria and the likelihood ratio tests of all the models with ARMA($p$,$q$) errors are summarised in table~\ref{ARMA} in the appendix.

@@

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and the coefficients are multiplicative. The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix. The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised. For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as a positive effect. If, for example, a marginal effect of an explanatory variable $x$ being equal to its standard deviation is $1.1$, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figure~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).

@@

In the following few paragraphs, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.

{
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} \\ Log of maize yield and weather, ARMA(1,1) errors}}
%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL % in the file KEN11d_finals2.R in folder RCodes\DecemberNew												
\label{MainEst} 
\begin{footnotesize}
\lineup
\begin{tabular}{@{}lllllll} 
\br 
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&\textbf{Estimate}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
    &\textbf{Estimate}&\textbf{F-value\tnote{a}}&\textbf{Estimate}&\textbf{F-value\tnote{a}}\\
\mr
\\
\vspace{-0.2cm}Intercept&$\m0.259^{***}$&$19.908$&$\m0.241^{*}$&$5.084$&$\m0.342^{**}$&$9.891$\\
  \\ \vspace{-0.2cm}Prec. total&$\m0.078^{*}$&$\05.358$&$\m0.007^{}$&$0.028$&$\m0.244^{***}$&$19.140$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$-0.028^{*}$&$\04.263$&$\m0.003$&$0.037$&$-0.127^{***}$&$23.384$\\
    \\ \vspace{-0.2cm}Prec. c. of var.&$-0.079^{\bullet}$&$\03.278$&$-0.031$ &$0.237$&$-0.097$&$\02.326$\\
  \\  \vspace{-0.2cm}Dry spell -length\tnote{b}&$-0.067^{*}$&$\04.800$&$-0.181^{**}$&$6.841$&$-0.012$&$\00.164$\\
  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.848$&$-0.156^{**}$&$8.032$&$-0.011$&$\00.087$\\
  \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.010$&$-0.214^{*}$&$5.409$&$-0.128$&$\01.525$\\
  \\  \vspace{-0.2cm}Temp. c. of var.&$\m0.042^{\bullet}$&$\02.961$&$\m0.032$&$0.355$&$\m0.058^{*}$&$\05.372$\\
  \\
  \hline
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.2cm}
\\
\hline
\\
  \vspace{-0.2cm}Intercept\\
 \\ 
 \hline
\vspace{-0.2cm} \\
\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
\vspace{-0.2cm}
\\  
\br
\end{tabular} 
\end{footnotesize}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
        \begin{adjustwidth}{1cm}{} 
    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
  \item[b] In number of days.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

@@

As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with increasing length of maximum dry spell linearly. This is in accordance with our expectations.

@@

With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its squared term. The F-statistic of the squared values of precipitation is $23.384$ and the F-value of the linear term of precipitation is $19.140$. The variable with the third highest value of the F-statistic is CV of temperature and the corresponding F-statistic is equal to $5.372$. Hence, the difference in importance of the precipitation and CV of temperature is relatively large. The standardised regression coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised regression coefficient of CV of temperature is $0.058$. In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of its coefficient is $-0.128$. However, average temperature is insignificant according to the Satterthwaite's p-value and also the value of its F-statistic is relatively low ($1.528$).

To sum up, temperature is much more important for the ASAL counties than for the non-ASAL counties where temperature is isignificant. On the other hand, precipitation is not significant for the ASAL counties while it is strongly significant for the non-ASAL counties. This is illustrated in figures~\ref{MarEff1}(c)-(f) in this section. The plot of the marginal effect of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(c) in this section) while it is hill-shaped and steep for the non-ASAL counties (figure \ref{MarEff1}(e) in this section) similarly as in the case of the whole sample. As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the both subsamples . Marginal effects of the remaining predictors are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.

@@

The AIC is an important criterion for comparison of competing models, however it does not provide any information about the absolute goodness of fit. In order to assess the absolute goodness of fit of a model, a different measure is needed. For a simple linear regression, the coeffiecient of determination $R^2$ or its adjusted version is typically used for this purpose. Nevertheless, obtaining the $R^2$ or the adjusted $R^2$ is not straightforward for linear mixed models or for models estimated by the maximum likelihood in general. In this study, we used the coefficient of determination of generalized linear mixed-effects models introduced by \cite{Nakagawa2013}. For each of our samples (all counties, ASAL and non-ASAL), we present two versions of the statistic in table~\ref{R2}. These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects. In our case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts. As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties,  about $13\%$ of the variance in the model based on the ASAL counties and about $15\%$ of the variance in the model based on the non-ASAL counties.

\begin{center}
\begin{threeparttable}
\caption{\label{R2}Coefficient of determination $R^2$ for mixed models}
\begin{indented}
\item[]
\begin{tabular}{@{}lcc}
\\[-1em]
\br
\\[-1em]
\textbf{Sample:}&Marginal\tnote{a}&Conditional\tnote{b}\\
\hline
{All counties}& $0.300$&$0.526$\\
{ASAL}& $0.134$&$0.353$\\
{non-ASAL}&  $0.153$&$0.436$\\
\br
\end{tabular}
\begin{tablenotes}
\begin{footnotesize}
    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
        \begin{adjustwidth}{1cm}{}
    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
     \end{adjustwidth}
     \end{footnotesize}
\end{tablenotes}

@@

Following a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity. With only two exceptions, the VIF is below $2$ for all predictors in our final models. For CV of precipitation, the value of VIF is $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it is equal to $2.182$ in the model based on the subsample of the non-ASAL counties. These values are well below the threshold of problematic multicollinearity. The other exception is the linear term of seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably due to presence of the squared term of precipitation. However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation is below $2$ in the other two models. Therefore, we can conclude that the models do not suffer from presence of multicollinearity.

@@

{\begin{threeparttable}
\normalsize
\caption{\normalsize{\textbf{ {Mixed  effects model:}} exponents of the coefficient estimates Log of maize yield and weather, ARMA(1,1) errors}}
%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL										
\label{KenARe11_exponents} 
\lineup
\begin{tabular}{@{}lllllll} 
\br \\
  \textbf{Fixed effects:}&\textit{\textbf{All counties}}&\textit{\textbf{ASAL}}&\textit{\textbf{non-ASAL}}\\
\mr
\\
\vspace{-0.2cm}Intercept&$1.296^{***}$&$1.272^{*}$&$1.408^{**}$\\
  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$1.007^{}$&$1.277^{***}$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$1.003$&$0.881^{***}$\\
    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$0.970$ &$0.907^{}$\\
  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
  \\ \vspace{-0.2cm}Temp. - average&$0.820^{***}$&$0.808^{*}$&$0.880$ $^{}$\\
  \\  \vspace{-0.2cm}Temp. c. of var.&$1.043^{\bullet}$&$1.032$&$1.060$ ${*}$\\
  \\
  \hline\\[-1em]
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
  \\[-1em]
\hline
\\[-1em]Intercept\\
 \\[-1em] \hline
\\[-1em]
\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}
\\
\br
\end{tabular} 
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

>>>>

To verify our results, we adapted an alternative step-down model building
approach using the Satterthwaite’s method to determine the p-values of the individual
t-tests \cite{lmerTest}. We started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the first part of the procedure, insignificant random effects were removed one by one. In each iteration, the insignificant variable with the highest p-value was removed. This was repeated until only significant random effects remained. After this, we applied an analogous procedure to the fixed effects until we were left only with significant variables. As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$ for fixed effects \cite{lmerTest}. This method led us to the same results as those obtained with our primary AIC-based method described above.

According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991; 1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$). To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with an ARMA($p$,$q$) error structure. In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$. Comparing the AIC criteria and using the corresponding likelihood ratio test statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$). The value of the AIC criterion of the model with the ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.  The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $0.003$, hence the ARMA($1$,$1$) error structure turned out to be a better fit. The AIC criteria and the likelihood ratio tests of all the models with ARMA($p$,$q$) errors are summarised in table~\ref{ARMA} in the appendix.

@@

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and the coefficients are multiplicative. The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix. The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised. For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as a positive effect. If, for example, a size of a marginal effect of an explanatory variable $x$ was equal to $1.1$ for $x$ being equal to its standard deviation, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figure~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).

@@

In the following few paragraphs we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.

{
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} \\ Log of maize yield and weather, ARMA(1,1) errors}}
%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL % in the file KEN11d_finals2.R in folder RCodes\DecemberNew												
\label{MainEst} 
\begin{footnotesize}
\lineup
\begin{tabular}{@{}lllllll} 
\br 
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&\textbf{Estimate}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
    &\textbf{Estimate}&\textbf{F-value\tnote{a}}&\textbf{Estimate}&\textbf{F-value\tnote{a}}\\
\mr
\\
\vspace{-0.2cm}Intercept&$\m0.259^{***}$&$19.908$&$\m0.241^{*}$&$5.084$&$\m0.342^{**}$&$9.891$\\
  \\ \vspace{-0.2cm}Prec. total&$\m0.078^{*}$&$\05.358$&$\m0.007^{}$&$0.028$&$\m0.244^{***}$&$19.140$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$-0.028^{*}$&$\04.263$&$\m0.003$&$0.037$&$-0.127^{***}$&$23.384$\\
    \\ \vspace{-0.2cm}Prec. c. of var.&$-0.079^{\bullet}$&$\03.278$&$-0.031$ &$0.237$&$-0.097$&$\02.326$\\
  \\  \vspace{-0.2cm}Dry spell -length\tnote{b}&$-0.067^{*}$&$\04.800$&$-0.181^{**}$&$6.841$&$-0.012$&$\00.164$\\
  \\ \vspace{-0.2cm}Dry spells 	$\geq~4$ d.\tnote{c}&$-0.063^{*}$&$\04.848$&$-0.156^{**}$&$8.032$&$-0.011$&$\00.087$\\
  \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.010$&$-0.214^{*}$&$5.409$&$-0.128$&$\01.525$\\
  \\  \vspace{-0.2cm}Temp. c. of var.&$\m0.042^{\bullet}$&$\02.961$&$\m0.032$&$0.355$&$\m0.058^{*}$&$\05.372$\\
  \\
  \hline
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.2cm}
\\
\hline
\\
  \vspace{-0.2cm}Intercept\\
 \\ 
 \hline
\vspace{-0.2cm} \\
\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
\vspace{-0.2cm}
\\  
\br
\end{tabular} 
\end{footnotesize}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
        \begin{adjustwidth}{1cm}{} 
    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
  \item[b] Length of the longest dry spell in number of days.
    \item[c] Number of dry spells lasting for four days or longer. 
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

@@

As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with increasing length of maximum dry spell roughly linearly. This is in accordance with our expectations.

@@

With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its squared term. The F-statistic of the squared values of precipitation is $23.384$ and the F-value of the linear term of precipitation is $19.140$. The variable with the third highest value of the F-statistic is CV of temperature and the corresponding F-statistic is equal to $5.372$. Hence, the difference in importance of precipitation and CV of temperature is relatively large. The standardised regression coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised regression coefficient of CV of temperature is $0.058$. In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of its coefficient is $-0.128$. However, average temperature is insignificant according to the Satterthwaite's p-value $(0.217)$ and also the value of its F-statistic is relatively low ($1.528$).

To sum up, temperature is much more important for the ASAL counties than for the non-ASAL counties where temperature is isignificant. On the other hand, precipitation is not significant for the ASAL counties while it is strongly significant for the non-ASAL counties. This is illustrated in figures~\ref{MarEff1}(c)-(f) in this section. The plot of the marginal effect of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(c) in this section) while it is hill-shaped and steep for the non-ASAL counties (figure \ref{MarEff1}(e) in this section) similarly as in the case of the whole sample. As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the both subsamples. Marginal effects of the remaining predictors are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.

@@

The AIC is an important criterion for comparison of competing models, however it does not provide any information about the absolute goodness of fit. In order to assess the absolute goodness of fit of a model, a different measure is needed. For a simple linear regression, the coeffiecient of determination $R^2$ or its adjusted version is typically used for this purpose. Nevertheless, obtaining the $R^2$ or the adjusted $R^2$ is not straightforward for linear mixed models or for models estimated by the maximum likelihood in general. In this study, we used the coefficient of determination of generalized linear mixed-effects models introduced by \cite{Nakagawa2013}. For each of our samples (all counties, ASAL and non-ASAL), we present two versions of the generalised $R^2$ in table~\ref{R2}. These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects. In our case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts. As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties,  about $13\%$ of the variance in the model based on the ASAL counties and about $15\%$ of the variance in the model based on the non-ASAL counties.

\begin{center}
\begin{threeparttable}
\caption{\label{R2}Coefficient of determination $R^2$ for mixed models}
\begin{indented}
\item[]
\begin{tabular}{@{}lcc}
\\[-1em]
\br
\\[-1em]
\textbf{Sample:}&Marginal\tnote{a}&Conditional\tnote{b}\\
\hline
{All counties}& $0.300$&$0.526$\\
{ASAL}& $0.134$&$0.353$\\
{non-ASAL}&  $0.153$&$0.436$\\
\br
\end{tabular}
\begin{tablenotes}
\begin{footnotesize}
    \item \textit{Notes:} The coefficient of determination for generalised linear mixed-effects models \cite{Nakagawa2013}.  \\
        \begin{adjustwidth}{1cm}{}
    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
     \end{adjustwidth}
     \end{footnotesize}
\end{tablenotes}

@@

Following a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity. With only two exceptions, the VIF was below $2$ for all predictors in our final models. For CV of precipitation, the value of VIF was $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it was equal to $2.182$ in the model based on the subsample of the non-ASAL counties. These values are well below the threshold of problematic multicollinearity. The other exception was the linear term of seasonal precipitation as its VIF was equal to $2.407$ in the model for the non-ASAL counties, probably due to presence of its squared term in the model. However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation was below $2$ in the other two models. Therefore, we can conclude that the models do not suffer from presence of multicollinearity.

@@

{\begin{threeparttable}
\normalsize
\caption{\normalsize{\textbf{ {Mixed  effects model:}} exponents of the coefficient estimates Log of maize yield and weather, ARMA(1,1) errors}}
%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL										
\label{KenARe11_exponents} 
\lineup
\begin{tabular}{@{}lllllll} 
\br \\
  \textbf{Fixed effects:}&\textit{\textbf{All counties}}&\textit{\textbf{ASAL}}&\textit{\textbf{non-ASAL}}\\
\mr
\\
\vspace{-0.2cm}Intercept&$1.296^{***}$&$1.272^{*}$&$1.408^{**}$\\
  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$1.007^{}$&$1.277^{***}$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$1.003$&$0.881^{***}$\\
    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$0.970$ &$0.907^{}$\\
  \\  \vspace{-0.2cm}Dry spell -length\tnote{a}&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
  \\ \vspace{-0.2cm}Dry spells $\geq~4$ d.\tnote{b}&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
  \\ \vspace{-0.2cm}Temp. - average&$0.820^{***}$&$0.808^{*}$&$0.880$ $^{}$\\
  \\  \vspace{-0.2cm}Temp. c. of var.&$1.043^{\bullet}$&$1.032$&$1.060$ ${*}$\\
  \\
  \hline\\[-1em]
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
  \\[-1em]
\hline
\\[-1em]Intercept\\
 \\[-1em] \hline
\\[-1em]
\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}
\\
\br
\end{tabular} 
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
            \begin{adjustwidth}{1cm}{} 
  \item[a] Length of the longest dry spell in number of days.
    \item[b] Number of dry spells lasting for four days or longer. 
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
